---
title: "Atmospheric CO2"
author: "John W Smith"
date: "1/31/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Atmospheric CO2

Here we are going to be extracting atmospheric CO2 data to use as drivers for the DALECev model. 

We will start by reading in the data.

```{r}
## read in co2 data
atm_carbon <- read.csv('./monthly_in_situ_co2_mlo.csv', skip = 56)
## set column names 
colnames(atm_carbon) <- c('Year', 'Month', 'DateE', 'Date', 'CO2', 'CO2_SA', 'Fit', 'SA_Fit', 'CO2_fill', 'SA_fill')
## print first few observations
head(atm_carbon)
```

So our goal is to make these into a data frame with: date and CO2.

```{r}
extract_co2 <- function(data, year){
  if (year %% 4 == 0){
    ndays <- c(31, 29, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
  } else{
    ndays <- c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
  }
  co2_info <- matrix(NA, nrow = 0, ncol = 2)
  colnames(co2_info) <- c('Date', 'CO2')
  co2_info <- as.data.frame(co2_info)
  CO2 <- c()
  dates_co2 <- c()
  for (i in 1:12){
    index <- which(data$Year == year & data$Month == i)
    CO2_val <- data$CO2[index]
    CO2 <- c(CO2, rep(CO2_val, ndays[i]))
    Date_val <- paste0(paste0(paste0(i, '-'), paste0(1:ndays[i], '-')), year)
    dates_co2 <- c(dates_co2, Date_val)
  }
  return(list(CO2 = CO2, Dates = dates_co2))
}
```

This function will extract the values, replicate them on the daily level for each month, and return these values plus the dates.

```{r}
test_function <- extract_co2(atm_carbon, year = 2000)
plot(test_function$CO2)
length(test_function$CO2)
```

Everything here looks good to go. Now let's extract some CO2 data from a couple of years, store it, and save it. Let's do, say, 2014 - 2021. 

```{r}
years <- 2014:2021
atm_CO2 <- c()
dates_extract <- c()
for (i in years){
  fn_output <- extract_co2(data = atm_carbon, year = i)
  atm_CO2 <- c(atm_CO2, fn_output$CO2)
  dates_extract <- c(dates_extract, fn_output$Dates)
}

plot(atm_CO2)

df <- as.data.frame(cbind(atm_CO2, dates_extract))
```


