---
title: "TALL Carbon"
author: "John Smith"
date: "02/01/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(neonstore)
```

In this document, we will be creating a carbon budget for Talladega National Forest using data from `NEON`, and the tools `neonstore` and `allodb`.

NEON data is organized by data product ID in the NEON Date Portal: <https://data.neonscience.org/static/browse.html>
 
You will be using the following data products
      - DP1.10098.001 (Woody plant vegetation structure)
      - DP1.00096.001 (Soil physical properties (Megapit))
      - DP1.00097.001 (Soil chemical properties (Megapit))

# Download data

First, we are going to want to define the site ID. The four letter site code for Talladega National Forest is `TALL`.
```{r}
site <- "TALL"
```

We will also define the data product ID that we are going to extract first (Woody plant vegetation structure).

```{r}
data_product <- "DP1.10098.001"
```

Second, we will use the `neon_download()` function from `neonstore` to create a database of files for our data product and site.
```{r}
neon_download(product = data_product, site = site)

print(unique(neon_index(site = site)$table))
```



# Calculate carbon in trees

In this section, we will be calculating carbon in live and dead trees at Talladega. At the end we will have a *site-level* mean carbon stock in *live trees* for each year that was sampled from the plots that are sampling the ecosystem under the flux tower (e.g., tower plots).  

Read in the separate tables

```{r}
## use neon_read to extract the desired data
ind_table <- neon_read('vst_apparentindividual', site = site)

map_tag_table <- neon_read('vst_mappingandtagging', site = site)

plot_table <- neon_read('vst_perplotperyear', site = site)

allometrics <- read_csv("Allometrics.csv") %>% 
  mutate(SPECIES = ifelse(SPECIES == "spp.", "sp.", SPECIES))

## link allodb

genus_species <- unlist(str_split_fixed(map_tag_table$scientificName, " ", 3))

map_tag_table <- map_tag_table %>% 
  mutate(GENUS = genus_species[,1], 
         SPECIES = genus_species[,2]) 

map_tag_table <- left_join(map_tag_table, allometrics, by = c("GENUS","SPECIES"))

map_tag_table <- map_tag_table %>% 
  mutate(B1 = ifelse(is.na(B1), -2.0127, B1),
         B2 = ifelse(is.na(B2), 2.4835, B2))

```

Select the key variables in each table and join into the individual table, making sure that we have `GENUS`, `SPECIES`, and `measurementHeight`. These will be important when we use `allodb` to estimate the amount of carbon.

```{r}

select_map_tag_table <- map_tag_table %>% 
  select(individualID, scientificName, B1, B2, GENUS, SPECIES) 

select_plot_table <- plot_table %>% 
  select(plotID,totalSampledAreaTrees,plotType) %>% 
  distinct(plotID, .keep_all = TRUE)

select_ind_table <- ind_table %>% 
  select(individualID, plotID, date, stemDiameter,plantStatus, measurementHeight) 

combined_table <- inner_join(select_ind_table, select_map_tag_table, by = "individualID") %>% 
  arrange(plotID,individualID)

combined_table <- inner_join(combined_table, select_plot_table, by = c("plotID")) %>% 
  arrange(individualID)

combined_table_dbh <- combined_table[which(combined_table$measurementHeight == 130),]

combined_table_dbh <- combined_table_dbh[which(!is.na(combined_table_dbh$stemDiameter)),]

```

Tidy up the individual tree data to include only live trees from the tower plots.  Also create a variable that is the year of the sample date.

```{r}
combined_table_live_tower <- combined_table_dbh %>% 
  mutate(year = year(date)) %>%
  filter(str_detect(plantStatus,"Live"),
         plotType == "tower") 
```

Calculate the biomass of each tree in the table. We will do this using `get_biomass` in the `allodb` package. This function takes as arguments: `dbh`, `genus`, `species`, `coords`. We have already extracted genus and species, and in fact we have already extracted dbh as well. The steps where we were subsetting based on measurement heights of 130 cm was actually subsetting to include only data that had dbh measurements. 

In this next section, as well as a future one where we calculate dead tree carbon, we are going to make a simplfying assumption. We will assume that the below ground biomass of a tree is some fixed proportion of its above ground biomass. In our analysis, we will assume this value is $.3$, but it is a parameter that can be changed.

```{r}
ag_bg_prop <- .3
```

```{r}
library(allodb)

tree_live_carbon <- combined_table_live_tower %>%
        mutate(ag_tree_kg = get_biomass(
          dbh = combined_table_live_tower$stemDiameter,
          genus = combined_table_live_tower$GENUS,
          species = combined_table_live_tower$SPECIES,
          coords = c(-87.3, 34.2)
          ),
         bg_tree_kg = ag_tree_kg * ag_bg_prop, ## assumption about ag to bg biomass
         tree_kgC = (ag_tree_kg + bg_tree_kg) * 0.5) ## convert biomass to carbon
```

Calculate the plot level biomass

```{r}
 plot_live_carbon <-  tree_live_carbon %>%
    mutate(treeC_kgCm2 = (tree_kgC)/(totalSampledAreaTrees))  %>%
    group_by(plotID, year) %>%
    summarise(plot_kgCm2 = sum(treeC_kgCm2, na.rm = TRUE))
```

```{r}
ggplot(plot_live_carbon, aes(x = year, y = plot_kgCm2, color = plotID)) + 
  geom_point() +
  geom_line()

```

```{r}
 site_live_carbon <- plot_live_carbon %>%
    group_by(year) %>%
    summarize(mean_kgCperm2 = mean(plot_kgCm2, na.rm = TRUE),
              sd_kgCperm2 = sd(plot_kgCm2))
```

```{r}
ggplot(site_live_carbon, aes(x = year, y = mean_kgCperm2)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=mean_kgCperm2-sd_kgCperm2, ymax=mean_kgCperm2+sd_kgCperm2), width=.2,
                 position=position_dodge(0.05))

```

# Calculate carbon in dead trees

```{r}
combined_table_dead_tower <- combined_table_dbh %>% 
  mutate(year = year(date)) %>%
  filter(grepl("Standing dead",plantStatus),
         plotType == "tower") 
```

Calculate the biomass of each tree in the table

```{r}
tree_dead_carbon <- combined_table_dead_tower %>%
    mutate(ag_tree_kg = get_biomass(
          dbh = combined_table_dead_tower$stemDiameter,
          genus = combined_table_dead_tower$GENUS,
          species = combined_table_dead_tower$SPECIES,
          coords = c(-87.3, 34.2)
          ),
         bg_tree_kg = ag_tree_kg * ag_bg_prop,
         tree_kgC = (ag_tree_kg + bg_tree_kg) * 0.5)
```

Calculate the plot level biomass

```{r}
 plot_dead_carbon <-  tree_dead_carbon %>%
    mutate(treeC_kgCm2 = (tree_kgC)/(totalSampledAreaTrees))  %>%
    group_by(plotID, year) %>%
    summarise(plot_kgCm2 = sum(treeC_kgCm2, na.rm = TRUE))
```

```{r}
 site_dead_carbon <- plot_dead_carbon %>%
    group_by(year) %>%
    summarize(mean_kgCperm2 = mean(plot_kgCm2, na.rm = TRUE),
              sd_kgCperm2 = sd(plot_kgCm2))
```

```{r}
ggplot(plot_dead_carbon, aes(x = year, y = plot_kgCm2, color = plotID)) + 
  geom_point() +
  geom_line()

```

# Calculate carbon in dead trees on the ground

```{r}
#Download bieogeochemistry soil data to get carbon concentration
data_product1 <- "DP1.10014.001"
#Download physical soil data to get bulk density
data_product2 <- "DP1.10010.001"
```

```{r}
site <- "TALL"
## use neon_download to get desired data products
neon_download(product = data_product1, site = site)

neon_download(product = data_product2, site = site)

print(unique(neon_index(site = site)$table))
```



```{r}
log_table <- neon_read("cdw_fieldtally", site = site)

log_bulk_density <- 0.03 #g per cm

#logMaxDiameter is in units of cm
#logLength is in units of m

log_table <- log_table %>% 
  mutate(year = year(date)) %>% 
  filter(plotType == "tower") %>% 
  mutate(CWD_kgC = 0.5 * logMaxDiameter * (logLength * 100) * 0.03 / 1000)

plot_cwd_carbon <- log_table %>% 
  group_by(plotID,year) %>% 
  summarise(plot_CWD_kgC = sum(CWD_kgC, na.rm = TRUE))

site_cwd_carbon <- plot_cwd_carbon %>% 
  group_by(year) %>% 
  summarize(mean_kgCperm2 = mean(plot_CWD_kgC),
            sd_CWD_kgC = sd(plot_CWD_kgC))
```

# Calculate carbon in soils

```{r}
#Download bieogeochemistry soil data to get carbon concentration
#data_product1 <- "DP1.00097.001"
#Download physical soil data to get bulk density
data_product2 <- "DP1.00096.001"
```

```{r}
## use neon_download to extract data product
neon_download(product = data_product2, site = site)
print(unique(neon_index(site = site)$table))
```
  
  
```{r}
mgc_perbiogeosample <- neon_read('mgp_perbiogeosample', site = site)
mgp_perbulksample <- neon_read('mgp_perbulksample', site = site)
```
  
```{r}  
bulk_density <- mgp_perbulksample %>% 
    filter(bulkDensSampleType == "Regular") %>% 
    select(horizonName,bulkDensExclCoarseFrag) 

  #gramsPerCubicCentimeter
horizon_carbon <- mgc_perbiogeosample %>% 
    filter(biogeoSampleType == "Regular") %>% 
    select(horizonName,biogeoTopDepth,biogeoBottomDepth,carbonTot) 
```
  

```{r}
  #Unit notes
  #bulkDensExclCoarseFrag = gramsPerCubicCentimeter
  #carbonTot = gramsPerKilogram
  
  #Combine and calculate the carbon of each horizon
horizon_combined <- inner_join(horizon_carbon,bulk_density, by = "horizonName") %>%
    #Convert volume in g per cm3 to mass per area in g per cm2 by multiplying by layer thickness
    mutate(horizon_soil_g_per_cm2 = (biogeoBottomDepth - biogeoTopDepth) * bulkDensExclCoarseFrag) %>% 
    #Units of carbon are g per Kg soil but we have bulk density in g per cm2 so convert Kg soil to g soil
    mutate(CTot_g_per_g_soil = carbonTot*(1/1000),  #Units are g C per g soil
           horizon_C_g_percm2 = CTot_g_per_g_soil*horizon_soil_g_per_cm2, #Units are g C per cm2
           horizon_C_kg_per_m2 = horizon_C_g_percm2 * 10000 / 1000) %>% #Units are g C per m2
    select(-CTot_g_per_g_soil,-horizon_C_g_percm2) %>%
    arrange(biogeoTopDepth)
  
site_soil_carbon <- horizon_combined %>% 
    summarize(soilC_gC_m2 = sum(horizon_C_kg_per_m2))
```

```{r}
ggplot(horizon_combined, map = aes(-biogeoTopDepth,horizon_C_kg_per_m2)) +
  geom_line() +
  geom_point() +
  labs(y = "Carbon", x = "Depth", title = "Soil carbon by depth") +
  coord_flip()
```

# Carbon Table

```{r}
total_carbon_components <- unlist(c(site_live_carbon$mean_kgCperm2[1],site_dead_carbon$mean_kgCperm2[1],site_cwd_carbon$mean_kgCperm2[1],site_soil_carbon)) 

total_carbon_components <- c(total_carbon_components,sum(total_carbon_components))

names(total_carbon_components) <- c("Live Trees", "Standing Dead", "Down Coarse Wood", "Soil", "Total")

total_carbon_components

```

